Robust Estimation, Prediction and Control with \\ Linear Dynamics and Arbitrary Costs

### 2020-01-17 main.tex
Add parts on planning

The dynamics of the real system are described in continuous time. However, all the sensing and control performed on-board happen in discrete time, with time-step $\dd t$. For any variable $z$, we use the subscript notation to refer to these discrete times: $z_n = z(n\dd t)$.

@@

From now on, we consider that we have reached step $N$, and that a confidence region $\cC_{[N],\delta}$ was obtained with data $\cD_{[N]}$ collected so far. The following sections tackle the optimisation of the robust control objective:
\begin{equation}
\label{eq:robust-control}
\max_\bu \underbrace{\min_{\substack{A(\theta) \in \cC_{[N],\delta} \\ \bw\in[\underline{\bw},\overline{\bw}]^\Natural}} \left[\sum_{n=1}^\infty \gamma^n R(x_n(\bu,\bw))\right]}_{V^r(\bu)}
\end{equation}
where $\bu = (u_{N+1},u_{N+2},\cdots)$ and $\bw = (\omega_N, \omega_{N+1},\cdots)$ are the sequences of future controls and perturbations, respectively. Here, we considered the bounded noise setting of \autoref{assumpt:bounded-noise}. In the sub-Gaussian noise setting of \autoref{assumpt:gaussian-noise}, we can use the same formulation by deriving high-confidence bounds $\underline{\omega}(t) \leq \omega(t) \leq \overline{\omega}(t)$ that each hold with probability $\delta_t$. By choosing $\delta_t = \frac{\delta}{t(t+1)}$, we have that the event $\{\forall t, \underline{\omega}(t) \leq \omega(t) \leq \overline{\omega}(t)\}$ holds with probability $1-\delta$.

@@

The objective of this section, as illustrated in \autoref{fig:interval-hull} is to  derive a \emph{set predictor} for the system \eqref{eq:dynamics}, which takes the information on the observed current state ${x}({0})$, some admissible bounds on the perturbation $[\underline{\omega}(t),\overline{\omega}(t)]$, and a confidence region $\cC_\delta$ in the form \eqref{eq:generic-polytope} containing $A(\theta)$; and verifies the inclusion property:
\begin{equation}
\label{eq:interval_property}
\underline{x}(t)\leq x(t)\leq\overline{x}(t),\quad\forall t\geq0.
\end{equation}

@@

\begin{theorem}[Interval Predictor \citep{leurent2019interval}]
\label{thm:predictor}
Assuming that \eqref{eq:polytope} is satisfied for the system \eqref{eq:dynamics}, then the interval predictor:
\begin{eqnarray}
\dot{\underline{x}}(t) & = & A_{0}\underline{x}(t)-\Delta A_{+}\underline{x}^{-}(t)-\Delta A_{-}\overline{x}^{+}(t)\nonumber \\
 &  & +Bu(t)+D^{+}\underline{\omega}(t)-D^{-}\overline{\omega}(t),\nonumber\\
\dot{\overline{x}}(t) & = & A_{0}\overline{x}(t)+\Delta A_{+}\overline{x}^{+}(t)+\Delta A_{-}\underline{x}^{-}(t) \label{eq:interval_predictor} \\
 &  & +Bu(t)+D^{+}\overline{\omega}(t)-D^{-}\underline{\omega}(t),\nonumber \\
 &  & \underline{x}(0)=\underline{x}_{0},\;\overline{x}(0)=\overline{x}_{0}\nonumber 
\end{eqnarray}
ensures the inclusion property \eqref{eq:interval_property}.

@@

\begin{align}
\underline{R}(t,\bu) \eqdef \min_{x\in[\underline{x}(t, \bu), \overline{x}(t, \bu)]}  R(x) \\
\hat{V}^r(\bu) \eqdef \sum_{t=0}^\infty \gamma^t \underline{R}(t,\bu) \label{eq:surrogate_objective}
\end{align}
where $\underline{x}(t, \bu), \overline{x}(t, \bu)$ follow the dynamics defined in \eqref{eq:interval_predictor}.
\end{definition}

This does not amount to defining a pessimistic reward-function $\underline{R}$, since the worst case is assessed over the whole trajectory being optimized, and not at the level of a local transition. In particular, it cannot be precomputed prior to the control optimization procedure, it needs to be done jointly.

%\begin{algorithm}[tp]
%  \SetAlgoLined\DontPrintSemicolon
%  \SetKwFunction{algo}{robust\_control}
%  \SetKwFunction{proc}{pessimistic\_step}
%  \SetKwProg{myalg}{Algorithm}{}{}
%  \myalg{\algo{$s_0$}}{
%  \While{resources available}
%  {
%  Select a candidate action sequence $a$ with a planning algorithm.\;
%  Sample a pessimistic transition and reward $\underline{r}_t$ using \proc{$\underline{x}(t)$, $\overline{x}(t)$, a}.
% \;
% Update the planning algorithm with the sampled pessimistic transition.
%  }
%  \KwRet $\argmax_{\bu} \hat{v^r}(\bu)$ \;}{}
%  \setcounter{AlgoLine}{0}
%  \SetKwProg{myproc}{Procedure}{}{}
%  \myproc{\proc{$\underline{x}(t)$, $\overline{x}(t)$, $a$}}{
%  Integrate \eqref{eq:interval_predictor} over $dt$ with controls $a$ to compute $\underline{x}(t+dt), \overline{x}(t+dt)$\;
%  Compute worst-case reward $\underline{r}_t = \min_{x\in[\underline{x}(t), \overline{x}(t)]} r(x)$\;
%  \KwRet $\underline{x}(t+dt), \overline{x}(t+dt), \underline{r}_t$\;}
%\caption{Interval-based Robust Control}
%\label{algo:irc}
%
%\end{algorithm}

@@

@@

\begin{definition}[Robust sequence upper bounds] We define an upper-bound for the value of a sequences of action $a$:
\begin{equation}
\label{eq:robust_sequence_ucb}
B_a^r(n)  \eqdef
\begin{cases}
\min_{m\in[M]} \sum_{t=0}^{h-1} \gamma^t R_t^m  + \frac{\gamma^h}{1-\gamma} &\text{if } a \text{ is a leaf;}\\
\max_{b\in\mathcal{A}} B_{ab}^r(n) & \text{else.}
\end{cases}
\end{equation}
where $R_t^m$ is the reward obtained at step $t$ by following $a$ with dynamics $f^m$.
\end{definition}
An illustration of the computation of the robust b-values is presented in Figure \ref{fig:drop}.

@@

Using the Cauchy-Schwartz inequality, we get:
\begin{align*}
    |x^\transp\theta_{N,\lambda}  -x^\transp\theta| &\leq \|x\|_{G_{N, \lambda}^{-1}}\left(\left\|\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n\right\|_{G_{N, \lambda}^{-1}} + \lambda\|\theta\|_{G_{N, \lambda}^{-1}}\right)
\end{align*}

@@

>>>>

The dynamics of the real system are described in continuous time. However, all the sensing and control performed on-board happen in discrete time, with time-step $\tau$. For any variable $z$, we use the subscript notation to refer to these discrete times: $z_n = z(n\tau)$.

@@

From now on, we consider that we have reached the time step $n=N$, and that a confidence region $\cC_{[N],\delta}$ was obtained with data $\cD_{[N]}$ collected so far. The following sections tackle the optimisation of the robust control objective:
\begin{equation}
\label{eq:robust-control}
\max_\bu \underbrace{\min_{\substack{A(\theta) \in \cC_{[N],\delta} \\ \bw\in[\underline{\bw},\overline{\bw}]^\Natural}} \left[\sum_{n=N+1}^\infty \gamma^n R(x_n(\bu,\bw))\right]}_{V^r(\bu)}
\end{equation}
where $\bu = (u_{N+1},u_{N+2},\cdots)$ and $\bw = (\omega_N, \omega_{N+1},\cdots)$ are the sequences of future controls and perturbations, respectively. Here, we considered the bounded noise setting of \autoref{assumpt:bounded-noise}. In the sub-Gaussian noise setting of \autoref{assumpt:gaussian-noise}, we can use the same formulation by deriving high-confidence bounds $\underline{\omega}(t) \leq \omega(t) \leq \overline{\omega}(t)$ that each hold with probability $\delta_t$. By choosing $\delta_t = \frac{\delta}{t(t+1)}$, we have that the event $\{\forall t, \underline{\omega}(t) \leq \omega(t) \leq \overline{\omega}(t)\}$ holds with probability $1-\delta$.

@@

The objective of this section, as illustrated in \autoref{fig:interval-hull} is to  derive a \emph{set predictor} for the system \eqref{eq:dynamics}, which takes the information on the observed current state ${x}({0})$, some admissible bounds on the perturbation $[\underline{\omega}(t),\overline{\omega}(t)]$, and a confidence region $\cC_\delta$ in the form \eqref{eq:generic-polytope} containing $A(\theta)$; and verifies the inclusion property:
\begin{equation}
\label{eq:interval_property}
\underline{x}(t)\leq x(t)\leq\overline{x}(t),\quad\forall t\geq N\tau.
\end{equation}

@@

\begin{theorem}[Interval Predictor \citep{leurent2019interval}]
\label{thm:predictor}
Assuming that \eqref{eq:polytope} is satisfied for the system \eqref{eq:dynamics}, then the interval predictor:
\begin{eqnarray}
\dot{\underline{x}}(t) & = & A_{0}\underline{x}(t)-\Delta A_{+}\underline{x}^{-}(t)-\Delta A_{-}\overline{x}^{+}(t)\nonumber \\
 &  & +Bu(t)+D^{+}\underline{\omega}(t)-D^{-}\overline{\omega}(t),\nonumber\\
\dot{\overline{x}}(t) & = & A_{0}\overline{x}(t)+\Delta A_{+}\overline{x}^{+}(t)+\Delta A_{-}\underline{x}^{-}(t) \label{eq:interval_predictor} \\
 &  & +Bu(t)+D^{+}\overline{\omega}(t)-D^{-}\underline{\omega}(t),\nonumber \\
 &  & \underline{x}(N\tau)=\overline{x}(N\tau)={x}(N\tau)\nonumber 
\end{eqnarray}
ensures the inclusion property \eqref{eq:interval_property}.

@@

\begin{align}
\underline{R}_n(\bu) \eqdef \min_{x\in[\underline{x}_n(\bu), \overline{x}_n(\bu)]}  R(x) \label{eq:pessimistic-rewards} \\
\hat{V}^r(\bu) \eqdef \sum_{n=N+1}^\infty \gamma^n \underline{R}_n(\bu) \label{eq:surrogate_objective}
\end{align}
where $\underline{x}_n(\bu), \overline{x}_n(\bu)$ follow the dynamics defined in \eqref{eq:interval_predictor}.
\end{definition}

This does not simply amounts to a change of reward function, since the worst case is assessed over the whole past trajectory being optimized, which means that this pessimistic reward $\underline{R}$ does not respect the Markov assumption.

@@

In order to optimize our surrogate objective $\hat{V}^r$, we turn to sequential planning algorithms. We cannot use the Dynamic Programming techniques because the state space is continuous and the pessimistic rewards are non-markovian. Instead, we turn to tree-based planning algorithms which optimize a sequence of actions based on the corresponding sequence of rewards without enumerating the states nor requiring the markovity of the rewards. In particular, we use the \texttt{OPD} algorithm \citep{Hren2008} tailored for the case where the sequence of rewards is a deterministic function of the sequence of actions, which is appropriate since both the predictor dynamics \eqref{eq:interval_predictor} and the corresponding pessimistic rewards \eqref{eq:pessimistic-rewards} are deterministic.

\texttt{OPD} progressively builds a tree $\cT_k$ by expanding the leaf $a_k\in\cL_k$ with highest B-value:
\begin{equation}
\argmax_{a\in\cL_k} B_a(k), \quad B_a(k) = \sum_{n=0}^{h-1} \underline{R}_n(a) + \frac{\gamma^h}{A-\gamma}.
\end{equation}

@@

\begin{definition}[Robust sequence upper bounds] We define an upper-bound for the value of a sequences of action $a$:
\begin{equation}
\label{eq:robust-b-values}
B_a^r(k)  \eqdef
\begin{cases}
\min_{m\in[M]} \sum_{n=0}^{h-1} \gamma^n R_n^m  + \frac{\gamma^h}{1-\gamma} &\text{if } a \text{ is a leaf;}\\
\max_{b\in\mathcal{A}} B_{ab}^r(k) & \text{else.}
\end{cases}
\end{equation}
where $R_t^m$ is the reward obtained at step $t$ by following $a$ with dynamics $f^m$.
\end{definition}
An illustration of the computation of the robust b-values is presented in Figure \ref{fig:drop}.

@@

Using the Cauchy-Schwartz inequality, we get:
\begin{align*}
    |x^\transp\theta_{N,\lambda}  -x^\transp\theta| \leq {} & \|x\|_{G_{N, \lambda}^{-1}}\left(\left\|\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n\right\|_{G_{N, \lambda}^{-1}}\right.\\ 
    &+ \left.\lambda\|\theta\|_{G_{N, \lambda}^{-1}}\right)
\end{align*}

@@

\subsection{Proof of \autoref{theorem:drop-regret}}

We start by showing the following lemma:

\begin{lemma}[Robust values ordering]
	In addition to the robust B-value defined in \eqref{eq:robust-b-values}, we also define the robust value of a sequence of actions $a$
	\begin{equation}
	V_a^r \eqdef \max_{\bu \in a\mathcal{A^\infty}} \min_{m\in[1, M]} \sum_{n=h(a)+1}^\infty \gamma^n R^m_n
	\end{equation}
	and the robust U-values of a sequence of action $a$
	\begin{equation}
	U_a^r(T)  \eqdef
	\begin{cases}
	\min_{m\in[M]} \sum_{n=0}^{h-1} \gamma^n R_n^m &\text{if } a \text{ is a leaf;}\\
	\max_{b\in\mathcal{A}} U_{ab}^r(n) & \text{else.}
	\end{cases}
	\end{equation}.

	Then, the robust values, U-values and B-values exhibit similar properties as the optimal values, U-values and B-values, that is: for all $0 < t < T$ and $a\in\mathcal{T}_T$,
	\begin{equation}
	U^r_a(t) \leq U^r_a(T) \leq V^r_a \leq B^r_a(T) \leq B^r_a(t)
	\end{equation}
	\label{lemma:uvb}
\end{lemma}
\begin{proof}
	By definition, when starting with sequence $i$, the value $u_i^m(n)$ represents the minimum admissible reward, while $b_i^m(n)$ corresponds to the best admissible reward achievable with respect to the the possible continuations of $i$. Thus, for all $i\in\mathcal{A}^*$, $u_i^m(n)$ and $u_i^r(n)$ are non-decreasing functions of $n$ and $b_i^m(n)$ and $b_i^r(n)$ are a non-increasing functions of $n$, while $v_i^m$ and $v_i^r$ do not depend on $n$.

	Moreover, since the reward function $r$ is assumed to have values in $[0, 1]$, the sum of discounted rewards from a node of depth $d$ is at most $\gamma^d + \gamma^{d+1}+\cdots = \frac{\gamma^d}{1-\gamma}$. As a consequence, for all $n \geq 0$ , $i\in\mathcal{L}_n$ of depth $d$, and any sequence of rewards $(r_t)_{t\in\mathbb{N}}$ obtained from following a path in $i\mathcal{A}^\infty$ with any dynamics $m \in [1, M]$:

	\begin{equation*}
	\sum_{t=0}^{d-1} \gamma^t r_t \leq \sum_{t=0}^{d-1} \gamma^t r_t + \sum_{t=d}^\infty \gamma^t r_t \leq \sum_{t=0}^{d-1} \gamma^t r_t + \frac{\gamma^d}{1-\gamma}
	\end{equation*}
	That is equivalent to:
	\begin{equation*}
	u^m_i(n) \leq \sum_{t=0}^\infty \gamma^t r_t \leq b^m_i(n) 
	\end{equation*}
	Hence,
	\begin{equation}
	\label{eq:min_m_values}
	\min_{m \in [1, M]} u^m_i(n) \leq \min_{m \in [1, M]} \sum_{t=0}^\infty \gamma^t r_t \leq \min_{m \in [1, M]} b^m_i(n)
	\end{equation}
	And as the left-hand and right-hand sides of \eqref{eq:min_m_values} are independent of the particular path that was followed in $i\mathcal{A}^\infty$, it also holds for the robust path:

	\begin{equation*}
	\min_{m \in [1, M]} u^m_i(n) \leq \max_{\pi\in i\mathcal{A}^\infty} \min_{m \in [1, M]} \sum_{t=0}^\infty \gamma^t r_t \leq \min_{m \in [1, M]} b^m_i(n)
	\end{equation*}
	that is,
	\begin{equation}
	\label{eq:urvrbr}
	u^r_i(n) \leq v^r_i  \leq b^r_i(n)
	\end{equation}

	Finally, \eqref{eq:urvrbr} is extended to the rest of $\mathcal{T}_n$ by recursive application of \eqref{eq:max_vr}, \eqref{eq:ur} and \eqref{eq:br}.
\end{proof}

We now turn to the proof of the theorem.

\begin{proof}
	\citet{Hren2008} first show in Theorem 2 that the simple regret of their optimistic planner is bounded by $\frac{\gamma^{d_n}}{1 - \gamma}$ where $d_n$ is the depth of $\mathcal{T}_n$. This properties relies on the fact that the returned action belongs to the deepest explored branch, which we can show likewise by contradiction using Lemma \ref{lemma:uvb}. This yields directly that $a = i_0$ where $i$ is some node of maximal depth $d_n$ expanded at round $t\leq n$, which by Algorithm \ref{algo:drop} verifies $b_a^r(t) = b_i^r(t) = \max_{x\in\mathcal{A}} b_x^r(t)$ and:
	\begin{align}
	\label{eq:Rndn}
	v^r - v_a^r &= v_{a^*}^r - v_a^r \\
	&\leq b_{a^*}^r(t) - v_a^r \\
	&\leq b_{a}^r(t) - u_a^r(t) \\
	&= b_{i}^r(t) - u_i^r(t) \\
	&= \frac{\gamma^{d_n}}{1-\gamma}.
	\end{align}

	Secondly, they bound the depth $d_n$ of $\mathcal{T}_n$ with respect to $n$. To that end, they show that the expanded nodes always belong to the sub-tree $\mathcal{T}_\infty$ of all the nodes of depth $d$ that are $\frac{\gamma^d}{1-\gamma}$-optimal. Indeed, if a node $i$ of depth $d$ is expanded at round $n$, then $b_i^r(n) \geq b_j^r(n)$ for all $j\in \mathcal{L}_n$ by Algorithm \ref{algo:drop}, thus the max-backups of \eqref{eq:br} up to the root yield $b^r_i(n) = b_\emptyset^r(n)$. Moreover, by Lemma \ref{lemma:uvb} we have that $b_\emptyset^r(n) \geq v_\emptyset^r = v^r$ and so $v_i^r \geq u_i^r(n) = b_i^r(n) - \frac{\gamma^d}{1-\gamma} \geq v^r - \frac{\gamma^d}{1-\gamma}$, thus $i \in \mathcal{T}_\infty$.

	Then from Assumption \ref{assumpt:beta} and the definition of $\beta$ applied to nodes in $\mathcal{T}_\infty$, there exists $d_0$ and $c$ such that the number $n_d$ of nodes of depth $d \geq d_0$ in $\mathcal{T}_\infty$ is bounded by $c\left(\frac{\gamma^d}{1-\gamma}\right)^\beta K^d$. As a consequence, 
	\begin{eqnarray*}
		n &= \sum_{d=0}^{d_n} n_d = n_0 + \sum_{d=d_0+1}^{d_n} n_d \\
		&\leq n_0 + \sum_{d=d_0+1}^{d_n} c\left(\frac{\gamma^d}{1-\gamma}\right)^\beta K^d \\
		&= n_0 + c'\sum_{d={d_0+1}}^{d_n} \kappa^d
	\end{eqnarray*}
	where $c'=\frac{c}{(1-\gamma)^\beta}$.

	\begin{itemize}
		\item If $\kappa > 1$, then $n \leq n_0 + c'\kappa^{d_0+1}\frac{\kappa^{d_n-d_0}-1}{\kappa-1}$ and thus $d_n \geq d_0 + \log_\kappa \frac{(n-n_0)(\kappa - 1)}{c'\kappa^{d_0+1}}$.
		We conclude from \eqref{eq:Rndn} that $\mathcal{R}_n \leq \frac{\gamma^{d_n}}{1-\gamma} = \frac{1}{1-\gamma} \left( \frac{(n-n_0)(\kappa - 1)}{c'\kappa^{d_0+1}} \right)^\frac{\log \gamma}{\log \kappa} = O\left(n^{-\frac{\log 1/\gamma}{\log \kappa}}\right)$.

		\item If $\kappa = 1$, then $n \leq n_0 + c'(d_n-d_0)$, hence from \eqref{eq:Rndn} we have $\mathcal{R}_n = O\left(\gamma^{nc'}\right)$.
	\end{itemize}
\end{proof}

\begin{proof}
	For any $\theta \in \Theta$, $t \in [0, H]$ and any trajectory $(s_0, \cdots, s_t)$ sampled from $\pi$ and $T_\theta$,

	\begin{equation*}
	s_t \in S(t, s_0, \pi) \subset \square S(t, s_0, \pi)
	\end{equation*}

	Hence, 
	\begin{equation*}
	R^{T_\theta}_\pi = \sum_{t=0}^\infty \gamma^t r(s_t, a_t) \geq \sum_{t=0}^H \gamma^t r(s_t, a_t) \geq \sum_{t=0}^H \min_{s\in \square S(t, s_0, \pi)} \gamma^t r(s, \pi(s)) = \hat{v^r}(\pi)
	\end{equation*}

	And finally,
	\begin{align*}
	v^r(\pi) = \min_{\theta\in\Theta} {v_\pi^{T_\theta}} = \min_{\theta\in\Theta} \expectedvalue(R^{T_\theta}_\pi) \geq \hat{v^r}(\pi)
	\end{align*}
\end{proof}
